---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(sf)
library(googlesheets4)
library(ggplot2)
library(lubridate)
library(there)
library(units)
library(ggpubr)
```

Loading in spatial data 
```{r}
settled_points_2018 <- st_read(
  dsn = here_file("Projects" , "Dispersal_paper", "settlement_designations"),
  layer = "settlement_locations_2018"
) %>% st_transform(st_crs(5070))

settled_points_2019 <- st_read(
  dsn = here_file("Projects" , "Dispersal_paper", "settlement_designations"),
  layer = "settlement_locations_2019"
) %>% st_transform(st_crs(5070))

settled_points <- rbind(settled_points_2018, settled_points_2019)

release_sites <- st_read(
  dsn = getwd(), 
  layer = "Release_sites") %>% 
  st_transform(st_crs(5070))

```

Loading, prepping, and joining the 2019 capture data
```{r}
banding_2019 <- read_sheet("1Ys9wnY8GV-kEO9q1hXWbvU3RGAzDe8L2lqWIX3wqQMw", "2019")
banding_2018 <- read_sheet("1Ys9wnY8GV-kEO9q1hXWbvU3RGAzDe8L2lqWIX3wqQMw", "2018")
banding_2017 <- read_sheet("1Ys9wnY8GV-kEO9q1hXWbvU3RGAzDe8L2lqWIX3wqQMw", "2017")
banding_2016 <- read_sheet("1Ys9wnY8GV-kEO9q1hXWbvU3RGAzDe8L2lqWIX3wqQMw", "2016")

banding <- rbind(banding_2016, banding_2017, banding_2018, banding_2019)

banding %>%
  filter(!is.na(BandNumber)) ->
  banding

names(settled_points)[1] <- "Haukos_SatID"

settled_points <-
  settled_points %>%
  left_join(banding)
```

Determining the distance between each point and the correct release site
```{r}
settled_points <- settled_points %>%
  mutate(distance_lasvacas = st_distance(settled_points, release_sites[2,])) %>%
  mutate(distance_p3 = st_distance(settled_points, release_sites[3,])) %>%
  mutate(distance_l7 = st_distance(settled_points, release_sites[4,])) %>%
  mutate(distance_l4 = st_distance(settled_points, release_sites[5,]))

correct_dist <- function(release, distance_lasvacas, distance_p3, distance_l7, distance_l4){
  switch (release,
    "Las Vacas Blancas" = distance_lasvacas,
    "P3" = distance_p3,
    "L7" = distance_l7,
    "L4" = distance_l4
  )
}

settled_points <- settled_points %>%
  rowwise() %>%
  mutate(dist_releasesite = correct_dist(ReleaseLocation, distance_lasvacas, distance_p3, distance_l7, distance_l4)) %>% 
  ungroup()

mean(settled_points$dist_releasesite/1000)
range(settled_points$dist_releasesite/1000)

#percent dispersed
settled_points %>%
  filter(as.numeric(dist_releasesite) > 5000) %>%
  nrow(.)/nrow(settled_points)
```

Figure: displacement from release site
```{r}
displacement_releasesite <- ggplot(data = settled_points, aes(fill = Sex)) +
  theme_bw() +
  geom_boxplot(aes(x = as_factor(ReleaseYear), y = as.numeric(dist_releasesite/1000))) +
  xlab("Year of release") +
  ylab("Release site to settlement point (km)") +
  theme(axis.title.x = element_blank())
  #ggtitle("Displacement from Release Site by Sex and Year") +
  #EnvStats::stat_n_text(aes(x = as_factor(ReleaseYear), y = as.numeric(dist_releasesite/1000)))

displacement_releasesite

#ggsave("displacement_releasesite_12_14_22.tiff",  width = 5, height = 3.5)
```


Calculating distance moved statistics
```{r}
dispersal_points_2018 <- st_read(dsn = here_file("Projects" , "Dispersal_paper", "settlement_designations"), layer = "dispersal_locations_2018")
dispersal_points_2019 <- st_read(dsn = here_file("Projects" , "Dispersal_paper", "settlement_designations"), layer = "dispersal_locations_2019")

dispersal_points <- rbind(dispersal_points_2018, dispersal_points_2019)

names(dispersal_points)[1] <- "Haukos_SatID"; names(dispersal_points)[2] <- "DateTime";  names(dispersal_points)[3] <- "reg_group"

dispersal_points %>%
  st_drop_geometry() %>%
  group_by(Haukos_SatID) %>%
  summarise(total_dist = sum(dist, na.rm = TRUE), last_disp_date = max(DateTime)) %>%
  left_join(banding) %>%
  mutate(dispersal_period = last_disp_date - as_date(HandlingDate)) ->
  dispersal_results

```

Figure: Distance moved before settlement
```{r}
distance_moved_before_settlement <-ggplot(data = dispersal_results, aes(fill = Sex)) +
  theme_bw() +
  geom_boxplot(aes(x = as_factor(ReleaseYear), y = total_dist/1000)) +
  xlab("Year of release") +
  ylab("Distance moved during dispersal (km)") +
  theme(axis.title.x = element_blank())
  #ggtitle("Distance Moved Before Settlement by Sex and Year") +
  #EnvStats::stat_n_text(aes(x = as_factor(ReleaseYear), y = total_dist/1000))

distance_moved_before_settlement

#ggsave("distance_moved_before_settlement_12_14_22.tiff",  width = 5, height = 3.5)
```

Figure: Days before settlement
```{r}
days_before_settlement <- ggplot(data = dispersal_results, aes(fill = Sex)) +
  geom_boxplot(aes(x = as_factor(ReleaseYear), y = dispersal_period)) +
  theme_bw() +
  xlab("Year of release") +
  ylab("Days before settlement")  +
  theme(axis.title.x = element_blank())
  #ggtitle("Days Moved Before Settlement by Sex and Year") +
  #EnvStats::stat_n_text(aes(x = as_factor(ReleaseYear), y = dispersal_period))

days_before_settlement

#ggsave("days_before_settlement_12_14_22.tiff", width = 5, height = 3.5)
```

Dispersal distance and variation stats
```{r}
mean(dispersal_results$total_dist/1000)
range(dispersal_results$total_dist/1000)
mean(dispersal_results$dispersal_period)

dispersal_results %>%
  filter(Sex == "Male") %>%
  pull(total_dist) %>%
  range()

dispersal_results %>%
  filter(Sex == "Female") %>%
  pull(dispersal_period) %>%
  sd()
```

Finally, calculating the # of leks visited during dispersal. Checked that this was working properly 12/14/22
```{r}
leks <- st_read(dsn = getwd(), "lek_sites") %>%
  st_transform(5070)

buffered_leks <- leks %>%
  st_buffer(500) %>% #3200
  dplyr::select(LekName)

num_leks_visited_df <- dispersal_points %>% 
  st_join(buffered_leks) %>% 
  st_drop_geometry() %>% 
  group_by(Haukos_SatID) %>% 
  summarise(n_leks_visited = length(na.omit(unique(LekName)))) %>% #leks_visited = na.omit(unique(LekName)),
  left_join(banding) %>% 
  mutate(ReleaseYear = as_factor(year(HandlingDate))) #%>% 
  #filter(ReleaseYear == "2018",
  #       Sex == "Male")

num_leks_visited <- ggplot(data = num_leks_visited_df, aes(fill = Sex)) +
  theme_bw() +
  geom_boxplot(aes(x = ReleaseYear, y = n_leks_visited)) +
  xlab("Year of Release") +
  ylab("Leks visited during dispersal") +
  theme(axis.title.x = element_blank())

num_leks_visited

num_leks_visited_df %>% 
  group_by(Sex) %>% 
  summarize(MEAN = mean(n_leks_visited), SD = sd(n_leks_visited), RANGE = c(range(n_leks_visited)))

#ggsave("num_leks_visited_12_14_22.tiff", width = 5, height = 3.5)
```

Creating a figure that incorporates all of these metrics
```{r fig.height=7, fig.width=7}
ggarrange(distance_moved_before_settlement, 
          displacement_releasesite, 
          days_before_settlement,
          disp_plot,
          #num_leks_visited, 
          common.legend = TRUE, 
          legend = "bottom", 
          labels = "AUTO")

ggsave("multipanel_dispersal_characteristics_1_6_23.tiff", height = 7.5, width = 8.6)
```

