---
title: "R Notebook"
output: html_notebook
---
Designating settlement sites to describe distances moved before settlement and displacement from the release site
Liam A. Berigan, 9/21/19

Table of Contents:
## 1: Creation and segmentation of trajectories, and designation of the response variable


Start by detaching all packages so I can customize load order
```{r}
library(readr)
library(sp) #base spatial packages for the analysis
library(sf) #easier to iterate over than standard spatial features
library(tidyverse) #mostly used for dplyr pipes and purrr's map function
library(adehabitatLT) #trajectory framework
library(googlesheets4)
library(lubridate)
library(there)
library(moveHMM)
```


#
## 1: Creation and segmentation of trajectories, and designation of the response variable
#

```{r}
lpc_gps <- read_csv(here_file("used_available_toolshare", "sat_locations_clean", "sat_locations_2019-11-10.csv"), col_types = cols(Haukos_SatID = col_character()))

lpc_gps$Haukos_SatID %>%
  str_pad(width = 11, side = "right", pad = "0") ->
  lpc_gps$Haukos_SatID
```

Bringing in additional data
Capture
```{r}
banding_2019 <- read_sheet("1Ys9wnY8GV-kEO9q1hXWbvU3RGAzDe8L2lqWIX3wqQMw", "2019")
banding_2018 <- read_sheet("1Ys9wnY8GV-kEO9q1hXWbvU3RGAzDe8L2lqWIX3wqQMw", "2018")
banding_2017 <- read_sheet("1Ys9wnY8GV-kEO9q1hXWbvU3RGAzDe8L2lqWIX3wqQMw", "2017")
banding_2016 <- read_sheet("1Ys9wnY8GV-kEO9q1hXWbvU3RGAzDe8L2lqWIX3wqQMw", "2016")

banding <- rbind(banding_2016, banding_2017, banding_2018, banding_2019)

banding %>%
  filter(!is.na(BandNumber)) ->
  banding
```

Mortality
```{r}
mort_ks <- read_sheet("1WnEf_bP3W_cDHcIVL6bBAnnBMkLRDlpEi0mamg_2TU8", "KS Mortality")
mort_co <- read_sheet("1WnEf_bP3W_cDHcIVL6bBAnnBMkLRDlpEi0mamg_2TU8", "CO Mortality")

mort <- rbind(mort_ks, mort_co)
```

Joining that data and filtering by mort date and release date
Removing all locations on and before the release date, and all locations on and after the mort date 
```{r}
lpc_gps %>%
  left_join(., dplyr::select(banding, Haukos_SatID, BandNumber, HandlingDate, ReleaseSite, ReleaseLocation, Sex, ReleaseYear)) %>%
  left_join(., dplyr::select(mort, BandNumber, Mortality_Detection_Date)) %>%
  dplyr::select(-surviving, -Point_ID) %>%
  mutate(Mortality_Detection_Date = replace_na(Mortality_Detection_Date, replace = "2021-1-1"))->
  lpc_gps

lpc_gps %>%
  mutate(Date = mdy(Date)) %>% 
  mutate(HandlingDate = as_date(HandlingDate),
         Mortality_Detection_Date = as_date(Mortality_Detection_Date)) %>%
  filter(Date > HandlingDate) %>%
  filter(Date < Mortality_Detection_Date) -> 
  lpc_gps
```

Only concerned with birds from release year 2019
```{r}
lpc_gps %>%
  filter(ReleaseYear == 2019) ->
  lpc_gps
```


Projecting, getting xy coordinates in 5070, and putting them back in the tibble
Also filtering out locations from August and after
```{r}
lpc_gps %>%
  filter(!is.na(Latitude)) %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) %>%
  st_transform(5070)->
  lpc_gps_sf

lpc_gps %>%
  filter(!is.na(Latitude)) %>%
  cbind(., st_coordinates(lpc_gps_sf)) %>%
  rename(ID = Haukos_SatID) %>%
  filter(Date < mdy("8/1/2019")) %>%
  mutate(DateTime = as.POSIXct(mdy_hm(DateTime))) ->
  lpc_gps


df_list <- lpc_gps %>% split.data.frame(.$ID)
```

Change the frequency of observations, only drawing one a day for each bird
I do this to create an average daily speed, which fits Markov models better when I'm concerned about patterns over larger time scales
```{r}
once_a_day <- function(i){
  i[!duplicated(i$Date),] %>% return()
}

df_list <- df_list %>% map(once_a_day)
```

Removing any trajectories that have fewer than 10 locations, as the algorithm won't be able to run on samples that small. Gets rid of 12 trajectories
```{r}
remove_small_samples <- function(i){
  indic <- nrow(i) > 9
  ifelse(indic, TRUE, FALSE) %>% return
}
valid_sample <- df_list %>% map(remove_small_samples) %>% unlist(use.names = FALSE)
df_list  <- df_list[valid_sample] #includes only adequate samples
```

Putting the trajectories into moveHMM's native trajectory format
```{r}
prepare_data <- function(i){
  i %>% prepData(type = "UTM", coordNames = c("X", "Y")) %>% return()
}
traj_list <- df_list %>% map(prepare_data)
```

Take out any trajectories which end before May 1st (early death). This changed from July 31.
I'm running an experiment and seeing what happens when I get rid of this
```{r}
#dater <- function(i){ #make a list of those which end at the correct date
#  indic <- any(i$Date > as.Date("2018/05/1"))
#  ifelse(indic, TRUE, FALSE) %>% return
#}
#correct_dates <- traj_list %>% map(dater) %>% unlist(use.names = FALSE)
#traj_list <- traj_list[correct_dates]#includes only correct dates
```

```{r}
hmm_fitter <- function(i){
  #First: using a moving window function to estimate parameters for the MLE
  dist_mean <- c()
  dist_sd <- c()
  rel_angle_mean <- c()
  
  for(t in 1:(nrow(i)-9)){
    moving_window <- i[t:(t+9),]
    moving_window %>%
      pull(step) %>%
      mean(na.rm = TRUE) %>%
      append(dist_mean, .) ->
      dist_mean
    moving_window %>%
      pull(step) %>%
      sd(na.rm = TRUE) %>%
      append(dist_sd, .) ->
      dist_sd
    moving_window %>%
      pull(angle) %>%
      abs() %>%
      mean(na.rm = TRUE) %>%
      append(rel_angle_mean, .) ->
      rel_angle_mean
  }
  moving_window_df <- cbind(dist_mean, dist_sd, rel_angle_mean) %>% as.data.frame()
  
  dist_mean_disp <- max(dist_mean, na.rm = TRUE)
  dist_mean_settl <- min(dist_mean, na.rm = TRUE)
  
  moving_window_df %>%
    filter(dist_mean == dist_mean_disp) ->
    moving_window_disp
  
  moving_window_df %>%
    filter(dist_mean == dist_mean_settl) ->
    moving_window_settl
  
  moving_window_disp <- moving_window_disp[1,]
  moving_window_settl <- moving_window_settl[1,]
  
  dist_sd_disp <- moving_window_disp %>%
    pull(dist_sd)
  
  dist_sd_settl <- moving_window_settl %>%
    pull(dist_sd)
  
  rel_angle_mean_disp <- moving_window_disp %>%
    pull(rel_angle_mean)
  
  rel_angle_mean_settl <- moving_window_settl %>%
    pull(rel_angle_mean)  
  
  if(is.nan(rel_angle_mean_settl)){
    rel_angle_mean_settl <- rel_angle_mean_disp
  }
  
  if(dist_mean_settl == 0) {
    dist_mean_settl <- 0.0001
  }
  
  if(dist_sd_settl == 0) {
    dist_sd_settl <- 0.0001
  }
  
  #Second: Using a MLE to estimate the movement mode likelihood at each step
  #Remember to go back and incorporate zero inflation as described in the vignette (if it's necessary after getting rid of mort locations)
  
  mu0 <- c(dist_mean_settl, dist_mean_disp) # step mean (two parameters: one for each state)
  
  sigma0 <- c(dist_sd_settl, dist_sd_disp) # step SD
  
  #Zero inflation code
  #I do not expect the proportion of zero steps to differ between movement modes
  #If there are zero steps, include zeromass0 as a step parameter. If not, don't
  whichzero <- which(i$step == 0)
  proportion_zero <- length(whichzero)/nrow(i)
  
  if(proportion_zero != 0) {
    zeromass0 <- c(proportion_zero, proportion_zero) # step zero-mass
    stepPar0 <- c(mu0, sigma0, zeromass0)
  } else {
    stepPar0 <- c(mu0, sigma0)
  }
  
  angleMean0 <- c(rel_angle_mean_settl, rel_angle_mean_disp) # angle mean
  
  kappa0 <- c(1,1) # angle concentration
  anglePar0 <- c(angleMean0, kappa0)
  
  #Bug testing code
  i$ID %>% unique() %>% paste0("ID: ", .) %>% print()
  #paste0("Dispersal Mean:", dist_mean_disp) %>% print()
  #paste0("Settled Mean:", dist_mean_settl) %>% print()
  #paste0("Dispersal SD:", dist_sd_disp) %>% print()
  #paste0("Settled SD:", dist_sd_settl) %>% print()
  #paste0("Dispersal Angle:", rel_angle_mean_disp) %>% print()
  #paste0("Settled Angle:", rel_angle_mean_settl) %>% print()
  #paste0("Proportion of zero steps:", proportion_zero) %>% print()
  
  ## call to fitting function
   fitHMM(data = i, nbStates=2, stepPar0 = stepPar0,
                           anglePar0 = anglePar0) %>%
     return()
}

movehmm_objects <- traj_list %>% map(hmm_fitter)
```

Find the first step where it's more likely that the bird is settled than dispersing

This is where shit gets weird. I'm going to use these model likelihoods to make a fake modpartltraj object, convert my existing trajectories to ltrajs, and then use adehabitat's partmod.ltraj function to find out where the ideal partition point is
```{r}
moveHMM_modpartltraj_convert <- function(i){
  i %>%
    stateProbs() %>% 
    as.data.frame() %>%
    rename(mod.1 = V1, mod.2 = V2) %>%
    as.matrix() ->
    t
  
  indiceNA <- c(1:nrow(t))
  
  class(t) <- "modpartltraj"
  #There should be no NAs in these model likelihoods, so I set all rows to nna
  attr(t, "nna.places") <- indiceNA 
  return(t)
}

movehmm_objects %>%
  map(moveHMM_modpartltraj_convert) ->
  modpartltraj_objects
```

Now converting to a ltraj
```{r}
fit_traj <- function(t){ #making a map-able function to fit a trajectory to each dataframe
  indiv <- t %>% 
    pull(ID) %>%
    unique() #creating the trajectory id
  txy <- cbind(t$x, t$y) #creating a dataframe of x and y coords for the traj
  as.ltraj(xy = txy, date = t$DateTime, id = indiv) %>% return() #create a trajectory and return to the list
}
ltraj_list <- traj_list %>% map(fit_traj)
```

For these trajectories, there is no single likely transition from one movement mode to another.
Many of these birds likely died in the middle of the dispersal movement (I'll make individual determinations later)

Trajectories that aren't fitting: 
2: all steps are dispersing 2019.172587
3: all steps are dispersing "2019.172591"
5: all steps are dispersing "2019.172598"
5: all steps are dispersing "2019.172600"
10: all steps are dispersing "2019.172615"
29: algorithim broke 2019.174272
30: all steps are dispersing "2019.174277"
31: delayed start to dispersal broke algorithm "2019.174280"

```{r}
ltraj_list[[2]] <- NULL; modpartltraj_objects[[2]] <- NULL
ltraj_list[[3]] <- NULL; modpartltraj_objects[[3]] <- NULL
ltraj_list[[5]] <- NULL; modpartltraj_objects[[5]] <- NULL
ltraj_list[[5]] <- NULL; modpartltraj_objects[[5]] <- NULL
ltraj_list[[10]] <- NULL; modpartltraj_objects[[10]] <- NULL
ltraj_list[[29]] <- NULL; modpartltraj_objects[[29]] <- NULL
ltraj_list[[30]] <- NULL; modpartltraj_objects[[30]] <- NULL
ltraj_list[[31]] <- NULL; modpartltraj_objects[[31]] <- NULL
```

Using adehabitat's partmod.ltraj function to find out where the ideal partition point is
```{r}
partition_traj <- function(i){
  print(i)
  names(ltraj_list[i]) %>% print()
  ltraj_list[[i]] %>%
    partmod.ltraj(tr = ., 
                  npart = 2, 
                  mods = modpartltraj_objects[[i]], 
                  na.manage = "locf")
}

partitioned_trajectories <- 1:length(ltraj_list) %>% map(partition_traj) 
```


Now that I have these trajectory segments, I use them to create a column for the response variable
All locations in the first segment: 0
The first location in the second segment: 1
The remainder of locations in the second segment: 2

A loop to turn these segments into a table of groups that I can use as dependent variables in a regression
NOTE: "All belonging to the second group" only returns the daily points

Start by defining empty vectors that I can append data to
```{r}
s_Sat_ID <- c(); s_DateTime = c(); s_reg_group = c(); s_x = c(); s_y = c(); s_dist = c()
```

```{r}
for (i in 1:length(partitioned_trajectories)){
  seg <- partitioned_trajectories[[i]]
  #Specify the first of the second segment
  s_DateTime <- seg$ltraj[[2]]$date[[1]] %>% append(s_DateTime,.)
  s_x <- seg$ltraj[[2]]$x[[1]] %>% append(s_x,.)
  s_y <- seg$ltraj[[2]]$y[[1]] %>% append(s_y,.)
  s_dist <- seg$ltraj[[2]]$dist[[1]] %>% append(s_dist,.)
  s_Sat_ID <- names(ltraj_list)[i] %>% append(s_Sat_ID,.)
  s_reg_group <- 1 %>% append(s_reg_group,.)
  
  #Code to add in everything that happens in the second segment, not including the first point
  second_segment <- seg$ltraj[[2]][2:nrow(seg$ltraj[[2]]),]
  s_DateTime <- second_segment$date %>% append(s_DateTime,.)
  s_x <- second_segment$x %>% append(s_x,.)
  s_y <- second_segment$y %>% append(s_y,.)
  s_dist <- second_segment$dist %>% append(s_dist,.)
  third_length <- second_segment$date %>% length
  s_Sat_ID <- names(ltraj_list)[i] %>% rep(times = third_length) %>% append(s_Sat_ID,.)
  s_reg_group <- 2 %>% rep(times = third_length) %>% append(s_reg_group,.)
  
  #Specify all belonging to the first segment
  s_DateTime <- seg$ltraj[[1]]$date %>% append(s_DateTime,.)
  s_x <- seg$ltraj[[1]]$x %>% append(s_x,.)
  s_y <- seg$ltraj[[1]]$y %>% append(s_y,.)
  s_dist <- seg$ltraj[[1]]$dist %>% append(s_dist,.)
  second_length <- seg$ltraj[[1]]$date %>% length
  s_Sat_ID <- names(ltraj_list)[i] %>% rep(times = second_length) %>% append(s_Sat_ID,.)
  s_reg_group <- 0 %>% rep(times = second_length) %>% append(s_reg_group,.)
  
  #Debugging code
  #print(i)
  #print(length(s_DateTime))
  #print(length(s_reg_group))
}
```

use all these vectors to create a dataframe 
```{r}
reg_groups <- tibble(Haukos_SatID = s_Sat_ID, DateTime = s_DateTime, x = s_x, y = s_y, reg_group = s_reg_group, dist = s_dist)
```

Removing all points from certain trajectories that broke because they were all dispersal points or broke for other reasons

```{r}
reg_groups %>%
  filter(Haukos_SatID != "2019.172604") %>% # broken due to long pause before dispersal
  filter(Haukos_SatID != "2019.172610") %>% # broken due to belated dispersal movement
  filter(Haukos_SatID != "2019.172597") %>% #all dispersal
  filter(Haukos_SatID != "2019.172602") %>% #didn't work
  filter(Haukos_SatID != "2019.172611") %>% #doesn't fit pattern
  filter(Haukos_SatID != "2019.172616") %>% #pattern is broken
  filter(Haukos_SatID != "2019.172636a") %>% #didn't work
  filter(Haukos_SatID != "2019.174255") %>% #all dispersal
  filter(Haukos_SatID != "2019.174256") %>% #pattern is broken
  filter(Haukos_SatID != "2019.174268") -> #all points are dispersing
  reg_groups
```


And writing settlement and dispersing points to shapefiles
Settlement points
```{r}
reg_groups %>%
  filter(reg_group == 1) %>%
  st_as_sf(coords = c("x", "y"), crs = 5070) %>%
  st_write(dsn = here_file("Projects", "Dispersal_corridors", "Dispersal_stats_redo"), layer = "settlement_locations_2019", driver = "ESRI Shapefile")
```

Dispersing points
```{r}
reg_groups %>%
  filter(reg_group == 0) %>%
  st_as_sf(coords = c("x", "y"), crs = 5070) %>%
  st_write(dsn = here_file("Projects", "Dispersal_corridors", "Dispersal_stats_redo"), layer = "dispersal_locations_2019", driver = "ESRI Shapefile")
```