---
title: "R Notebook"
output: html_notebook
---

NOTA BENE: Giving up on this for the foreseeable future- I don't think I can provide enough info for readers to simulate from this without a complete walkthrough. Will likely need to be a seperate paper.

```{r}
library(raster)
library(tidyverse)
library(googlesheets4)
library(sf)
library(there) #fixed; some code is used to it being broken
library(readr)
library(lubridate)
library(mapview)
library(amt)
library(adehabitatLT)
library(hms)
library(ggplot2)
library(parallel)
library(AICcmodavg)
library(glmmTMB)
library(broom.mixed)
```

Step length and turn angle distributions
```{r}
ssf_data_true <- ssf_data %>% 
  filter(case_ == TRUE)

ssf_data_true %>% 
  group_by("Sex") %>% 
  summarise(mn_sl = mean(sl_), #meters
            sd_sl = sd(sl_),
            mn_ta = mean(ta_), #radians
            sd_ta = sd(ta_))

sl_hist <- hist(ssf_data_true$sl_, breaks = 34, freq = FALSE)

data.frame(range = paste0(sl_hist$breaks[1:34], " - ", sl_hist$breaks[2:35]),
           density = sprintf("%.8f", sl_hist$density*100),
           counts = sl_hist$counts) %>% 
  write.csv("sl_dist.csv")

ta_hist <- hist(ssf_data$ta_, breaks = 33, freq = FALSE)

data.frame(range = paste0(ta_hist$breaks[1:32], " - ", ta_hist$breaks[2:33]),
           density = sprintf("%.8f", ta_hist$density*100),
           counts = ta_hist$counts) %>% 
  write.csv("ta_dist.csv")
```

Plot the histograms and save them
```{r}
ggplot(ssf_data_true, aes(x = sl_/1000)) +
  geom_histogram(breaks = sl_hist$breaks/1000, aes(y = after_stat(count)/nrow(ssf_data_true)), color = "black", fill = "grey") + #
  theme_bw() +
  ylab("Percent of steps") +
  xlab("Step lengths (km)") +
  scale_y_continuous(labels = scales::percent_format())

ggsave(filename = "sl_dist.png", height = 7/1.5, width = 10/1.5)

ggplot(ssf_data_true, aes(x = ta_)) +
  geom_histogram(breaks = ta_hist$breaks, aes(y = after_stat(count)/nrow(ssf_data_true)), color = "black", fill = "grey") + #
  theme_bw() +
  ylab("Percent of turn angles") +
  xlab("Turn angles (radians)") +
  scale_y_continuous(labels = scales::percent_format())

ggsave(filename = "ta_dist.png", height = 7/1.5, width = 10/1.5)
```


## Preparing candidate models using the untransformed data
```{r}
muff.candidate.set <- list()
muff.candidate.set[[1]] <- glmmTMB(case_ ~ -1 + (1|step_id_), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA))))

muff.candidate.set[[2]] <- glmmTMB(case_ ~ -1 + log_sl_ + (1|step_id_), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA))))

muff.candidate.set[[3]] <- glmmTMB(case_ ~ -1 + prop_path_crop + log_sl_ + (1|step_id_) + (0 + prop_path_crop_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[4]] <- glmmTMB(case_ ~ -1 + prop_path_shrub + log_sl_ + (1|step_id_) + (0 + prop_path_shrub_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[5]] <- glmmTMB(case_ ~ -1 + prop_path_dev_open + log_sl_ + (1|step_id_) + (0 + prop_path_dev_open_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[6]] <- glmmTMB(case_ ~ -1 + prop_path_grass + log_sl_ + (1|step_id_) + (0 + prop_path_grass_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[7]] <- glmmTMB(case_ ~ -1 + prop_path_crp + log_sl_ + (1|step_id_) + (0 + prop_path_crp_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[8]] <- glmmTMB(case_ ~ -1 + prop_path_forest + log_sl_ + (1|step_id_) + (0 + prop_path_forest_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[9]] <- glmmTMB(case_ ~ -1 + prop_path_barren + log_sl_ + (1|step_id_) + (0 + prop_path_barren_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[10]] <- glmmTMB(case_ ~ -1 + prop_path_water + log_sl_ + (1|step_id_) + (0 + prop_path_water_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[11]] <- glmmTMB(case_ ~ -1 + prop_path_wetland + log_sl_ + (1|step_id_) + (0 + prop_path_wetland_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[12]] <- glmmTMB(case_ ~ -1 + prop_path_dev_lmh + log_sl_ + (1|step_id_) + (0 + prop_path_dev_lmh_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[13]] <- glmmTMB(case_ ~ -1 + cross_secondary + log_sl_ + (1|step_id_) + (0 + cross_secondary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[14]] <- glmmTMB(case_ ~ -1 + cross_street + log_sl_ + (1|step_id_) + (0 + cross_street_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[15]] <- glmmTMB(case_ ~ -1 + cross_transmission_line + log_sl_ + (1|step_id_) + (0 + cross_transmission_line_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[16]] <- glmmTMB(case_ ~ -1 + log_distance_to_secondary + log_sl_ + (1|step_id_) + (0 + log_distance_to_secondary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

#Street is probably going to have to be excluded (not converging). Actually, it looks okay?
muff.candidate.set[[17]] <- glmmTMB(case_ ~ -1 + log_distance_to_street + log_sl_ + (1|step_id_) + (0 + log_distance_to_street_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[18]] <- glmmTMB(case_ ~ -1 + log_distance_to_well + log_sl_ + (1|step_id_) + (0 + log_distance_to_well_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[19]] <- glmmTMB(case_ ~ -1 + log_distance_to_transmission_line + log_sl_ + (1|step_id_) + (0 + log_distance_to_transmission_line_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[20]] <- glmmTMB(case_ ~ -1 + shrub_binary + log_sl_ + (1|step_id_) + (0 + shrub_binary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[21]] <- glmmTMB(case_ ~ -1 + dev_open_binary + log_sl_ + (1|step_id_) + (0 + dev_open_binary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[22]] <- glmmTMB(case_ ~ -1 + grass_binary + log_sl_ + (1|step_id_) + (0 + grass_binary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[23]] <- glmmTMB(case_ ~ -1 + crp_binary + log_sl_ + (1|step_id_) + (0 + crp_binary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))
```

## Fitting the candidate models

Set the variance as constant at 1000 and fit the model
```{r}
fit_glmm_constant_variance <- function(x){
  #Fix the standard deviation of the first random term, which is the (1|step_id_) component in the above model equation
  #set the variance to 1000 by setting its log (sd) to log(1000)
  x$parameters$theta[1] <- log(1e3)
  
  #We need to tell glmmTMB not to change the variance by setting it to NA:
  #now set when I define each individual function, so this isn't necessary here
  #x$mapArg <- list(theta=factor(c(NA)))
  
  #print so I know which model to attribute error code to
  x$fr %>% names() %>% print()
  
  #Then fit the model and look at the results:
  glmmTMB:::fitTMB(x) %>% 
    return()
}

muff.candidate.set %>%
  map(.f = fit_glmm_constant_variance) ->
  muff.model.set

#model_figure <- fit_glmm_constant_variance(model_figure)
```

```{r}
covariate_results <- map2(muff.model.set, 1:length(muff.model.set), function(x, y){
  broom::tidy(x) %>% 
    mutate(model_num = y, .before = effect)
}) %>% 
  bind_rows() %>% 
  mutate(estimate = sprintf("%.6f", estimate),
         std.error = sprintf("%.6f", std.error)) %>% 
  filter(effect == "fixed")

write.csv(covariate_results, "appendix_covariate_results.csv")
```

```{r}
muff.model.set[[23]]$call$formula
```

