---
title: "using the Muff et al. 2019 approach to step selection analysis"
output: html_notebook
---

Additionally depends on all packages in issf_glmtmb_prep.Rmd
```{r}
library(glmmTMB)
library(broom.mixed)
```

z transformations of all dependent variables, to ensure that I can compare their betas 
```{r}
ssf_data <- ssf_data %>%
  mutate(log_sl_z = (log_sl_ - mean(log_sl_))/sd(log_sl_),
         prop_path_crop_z = (prop_path_crop - mean(prop_path_crop))/sd(prop_path_crop),
         prop_path_shrub_z = (prop_path_shrub - mean(prop_path_shrub))/sd(prop_path_shrub),
         prop_path_dev_open_z = (prop_path_dev_open - mean(prop_path_dev_open))/sd(prop_path_dev_open),
         prop_path_grass_z = (prop_path_grass - mean(prop_path_grass))/sd(prop_path_grass),
         prop_path_crp_z = (prop_path_crp - mean(prop_path_crp))/sd(prop_path_crp),
         prop_path_forest_z = (prop_path_forest - mean(prop_path_forest))/sd(prop_path_forest),
         prop_path_barren_z = (prop_path_barren - mean(prop_path_barren))/sd(prop_path_barren),
         prop_path_water_z = (prop_path_water - mean(prop_path_water))/sd(prop_path_water),
         prop_path_wetland_z = (prop_path_wetland - mean(prop_path_wetland))/sd(prop_path_wetland),
         prop_path_dev_lmh_z = (prop_path_dev_lmh - mean(prop_path_dev_lmh))/sd(prop_path_dev_lmh),
         cross_secondary_z = (cross_secondary - mean(cross_secondary))/sd(cross_secondary),
         cross_street_z = (cross_street - mean(cross_street))/sd(cross_street),
         cross_transmission_line_z = (cross_transmission_line - mean(cross_transmission_line))/sd(cross_transmission_line),
         log_distance_to_secondary_z = (log_distance_to_secondary - mean(log_distance_to_secondary))/sd(log_distance_to_secondary),
         log_distance_to_street_z = (log_distance_to_street - mean(log_distance_to_street))/sd(log_distance_to_street),
         log_distance_to_well_z = (log_distance_to_well - mean(log_distance_to_well))/sd(log_distance_to_well),
         log_distance_to_transmission_line_z = (log_distance_to_transmission_line - mean(log_distance_to_transmission_line))/sd(log_distance_to_transmission_line),
         shrub_binary_z = (shrub_binary - mean(shrub_binary))/sd(shrub_binary),
         dev_open_binary_z = (dev_open_binary - mean(dev_open_binary))/sd(dev_open_binary),
         grass_binary_z = (grass_binary - mean(grass_binary))/sd(grass_binary),
         crp_binary_z = (crp_binary - mean(crp_binary))/sd(crp_binary))

```

## Preparing candidate models using the z-transformed data
```{r}
muff.candidate.set <- list()
muff.candidate.set[[1]] <- glmmTMB(case_ ~ -1 + (1|step_id_), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA))))

muff.candidate.set[[2]] <- glmmTMB(case_ ~ -1 + log_sl_z + (1|step_id_), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA))))

muff.candidate.set[[3]] <- glmmTMB(case_ ~ -1 + prop_path_crop_z + log_sl_z + (1|step_id_) + (0 + prop_path_crop_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[4]] <- glmmTMB(case_ ~ -1 + prop_path_shrub_z + log_sl_z + (1|step_id_) + (0 + prop_path_shrub_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[5]] <- glmmTMB(case_ ~ -1 + prop_path_dev_open_z + log_sl_z + (1|step_id_) + (0 + prop_path_dev_open_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[6]] <- glmmTMB(case_ ~ -1 + prop_path_grass_z + log_sl_z + (1|step_id_) + (0 + prop_path_grass_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[7]] <- glmmTMB(case_ ~ -1 + prop_path_crp_z + log_sl_z + (1|step_id_) + (0 + prop_path_crp_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[8]] <- glmmTMB(case_ ~ -1 + prop_path_forest_z + log_sl_z + (1|step_id_) + (0 + prop_path_forest_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[9]] <- glmmTMB(case_ ~ -1 + prop_path_barren_z + log_sl_z + (1|step_id_) + (0 + prop_path_barren_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[10]] <- glmmTMB(case_ ~ -1 + prop_path_water_z + log_sl_z + (1|step_id_) + (0 + prop_path_water_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[11]] <- glmmTMB(case_ ~ -1 + prop_path_wetland_z + log_sl_z + (1|step_id_) + (0 + prop_path_wetland_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[12]] <- glmmTMB(case_ ~ -1 + prop_path_dev_lmh_z + log_sl_z + (1|step_id_) + (0 + prop_path_dev_lmh_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[13]] <- glmmTMB(case_ ~ -1 + cross_secondary_z + log_sl_z + (1|step_id_) + (0 + cross_secondary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[14]] <- glmmTMB(case_ ~ -1 + cross_street_z + log_sl_z + (1|step_id_) + (0 + cross_street_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[15]] <- glmmTMB(case_ ~ -1 + cross_transmission_line_z + log_sl_z + (1|step_id_) + (0 + cross_transmission_line_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[16]] <- glmmTMB(case_ ~ -1 + log_distance_to_secondary_z + log_sl_z + (1|step_id_) + (0 + log_distance_to_secondary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

#Street is probably going to have to be excluded (not converging). Actually, it looks okay?
muff.candidate.set[[17]] <- glmmTMB(case_ ~ -1 + log_distance_to_street_z + log_sl_z + (1|step_id_) + (0 + log_distance_to_street_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[18]] <- glmmTMB(case_ ~ -1 + log_distance_to_well_z + log_sl_z + (1|step_id_) + (0 + log_distance_to_well_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[19]] <- glmmTMB(case_ ~ -1 + log_distance_to_transmission_line_z + log_sl_z + (1|step_id_) + (0 + log_distance_to_transmission_line_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[20]] <- glmmTMB(case_ ~ -1 + shrub_binary_z + log_sl_z + (1|step_id_) + (0 + shrub_binary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[21]] <- glmmTMB(case_ ~ -1 + dev_open_binary_z + log_sl_z + (1|step_id_) + (0 + dev_open_binary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[22]] <- glmmTMB(case_ ~ -1 + grass_binary_z + log_sl_z + (1|step_id_) + (0 + grass_binary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

muff.candidate.set[[23]] <- glmmTMB(case_ ~ -1 + crp_binary_z + log_sl_z + (1|step_id_) + (0 + crp_binary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

# Full model: doesn't converge
# muff.candidate.set[[24]] <- glmmTMB(case_ ~ -1 + prop_path_crop_z + prop_path_shrub_z + prop_path_dev_open_z + prop_path_grass_z + prop_path_crp_z + prop_path_forest_z + prop_path_barren_z + prop_path_water_z + prop_path_wetland_z + prop_path_dev_lmh_z + cross_secondary_z + cross_street_z + cross_transmission_line_z + log_distance_to_secondary_z + log_distance_to_street_z + log_distance_to_well_z + log_distance_to_transmission_line_z + shrub_binary_z + dev_open_binary_z + grass_binary_z + crp_binary_z + log_sl_z + (1|step_id_) + (0 + prop_path_crop_z|Haukos_SatID) + (0 + prop_path_shrub_z|Haukos_SatID) + (0 + prop_path_dev_open_z|Haukos_SatID) + (0 + prop_path_grass_z|Haukos_SatID) + (0 + prop_path_crp_z|Haukos_SatID) + (0 + prop_path_forest_z|Haukos_SatID) + (0 + prop_path_barren_z|Haukos_SatID) + (0 + prop_path_water_z|Haukos_SatID) + (0 + prop_path_wetland_z|Haukos_SatID) + (0 + prop_path_dev_lmh_z|Haukos_SatID) + (0 + cross_secondary_z|Haukos_SatID) + (0 + cross_street_z|Haukos_SatID) + (0 + cross_transmission_line_z|Haukos_SatID) + (0 + log_distance_to_secondary_z|Haukos_SatID) + (0 + log_distance_to_street_z|Haukos_SatID) + (0 + log_distance_to_well_z|Haukos_SatID) + (0 + log_distance_to_transmission_line_z|Haukos_SatID) + (0 + shrub_binary_z|Haukos_SatID) + (0 + dev_open_binary_z|Haukos_SatID) + (0 + grass_binary_z|Haukos_SatID) + (0 + crp_binary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1:21))))



# Use only for the figure: now no longer relevant
# model_figure <- glmmTMB(case_ ~ -1 + crp_binary + log_distance_to_street + log_sl_ + (1|step_id_) + (0 + crp_binary|Haukos_SatID) + (0 + log_distance_to_street|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1, 2))))

# Correlation coefficient of 0.62 between prop_crp and crp at endpoint; however, this makes crp at endpoint have an insignificant p value, so the two are definitely interacting in the model
# muff.candidate.set[[24]] <- glmmTMB(case_ ~ -1 + crp_binary_z + log_distance_to_street_z + prop_path_crp_z + log_sl_z + (1|step_id_) + (0 + crp_binary_z|Haukos_SatID) + (0 + log_distance_to_street_z|Haukos_SatID) + (0 + prop_path_crp_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1, 2, 3))))
```

Landcov originally looked like the following (model 13 below) but was split into binary variables due to extreme eigenvalues
muff.candidate.set[[13]] <- glmmTMB(case_ ~ -1 + landcov + log_sl_ + (1|step_id_) + (0 + landcov|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

## Fitting the candidate models

Set the variance as constant at 1000 and fit the model
```{r}
fit_glmm_constant_variance <- function(x){
  #Fix the standard deviation of the first random term, which is the (1|step_id_) component in the above model equation
  #set the variance to 1000 by setting its log (sd) to log(1000)
  x$parameters$theta[1] <- log(1e3)
  
  #We need to tell glmmTMB not to change the variance by setting it to NA:
  #now set when I define each individual function, so this isn't necessary here
  #x$mapArg <- list(theta=factor(c(NA)))
  
  #print so I know which model to attribute error code to
  x$fr %>% names() %>% print()
  
  #Then fit the model and look at the results:
  glmmTMB:::fitTMB(x) %>% 
    return()
}

muff.candidate.set %>%
  map(.f = fit_glmm_constant_variance) ->
  muff.model.set

#model_figure <- fit_glmm_constant_variance(model_figure)
```


## Comparing the candidate models using AIC
```{r}
muff.candidate.names <- c("null", "log_sl", "prop crop", "prop shrub", "prop dev open", "prop grass", "prop crp", "prop forest", "prop barren", "prop water", "prop wetland", "prop dev lmh", "cross secondary", "cross street", "cross transmission", "dist secondary", "dist street", "dist well", "dist transmission", "shrub_binary", "dev_open_binary", "grass_binary", "crp_binary")
aictab(cand.set = muff.model.set, modnames = muff.candidate.names)
```

12/13/22: All models seem to have converged correctly
```{r}
#crp_binary
muff.model.set[[23]] %>% summary() #LPCs are more likely to use a step that ends in CRP

#dist street
muff.model.set[[17]] %>% summary() #LPCs are more likely to use a step that ends far from a road

#prop crp
muff.model.set[[7]] %>% summary() # LPCs are more likely to use a step when a good portion of that step is CRP

#prop grass
muff.model.set[[6]] %>% summary() #LPCs are less likely to use a step that is largely composed of grassland
```

## Plotting data and resulting models

Plotting the real data for CRP binary
```{r}
ggplot(ssf_data, aes(x = crp_binary_z, colour= case_)) +
  geom_density()

ggplot(ssf_data, aes(y = crp_binary, x= case_)) +
  geom_boxplot()

ssf_data %>%
  as.data.frame() %>%
  count(case_, crp_binary)
```

Plotting the real data for dist street
```{r}
ggplot(ssf_data, aes(x = log_distance_to_street, colour= case_))+
  geom_density() #Might be a case to fit log_distance_to_street^2?

ggplot(ssf_data, aes(x = distance_to_street, colour= case_))+
  geom_density() #Might be a case to fit log_distance_to_street^2?

ggplot(ssf_data, aes(x = log_distance_to_street, colour= case_))+
  geom_density() #Might be a case to fit log_distance_to_street^2?

ggplot(ssf_data, aes(x = 1/log_distance_to_street, colour= case_))+
  geom_density() #Might be a case to fit log_distance_to_street^2?
```

Plotting the real data for log distance to well
```{r}
ggplot(ssf_data, aes(x = log_distance_to_well_z, colour= case_))+
  geom_density()

ggplot(ssf_data, aes(x = distance_to_well, colour= case_))+
  geom_density() #Might be a case to fit log_distance_to_street^2?
```


Plotting the real data for proportion of path as CRP
```{r}
ggplot(ssf_data, aes(x = prop_path_crp, colour= case_))+
  geom_density() #Might be a case to fit log_distance_to_street^2?
```

Parameters here that are useful:
prediction type link: based on the scale of the linear predictors
plogis changes that from a logit to a probability scale
re.form = NA says that we don't want random effects to be included in the prediction (population level)
se.fit = TRUE says that we want standard error to be fit for each of these points
```{r}
# outlen <- 500
# 
# street_range <- range(ssf_data$log_distance_to_street)
# sim.streets <- seq(from = street_range[1], to = street_range[2], length.out = outlen)
# 
# sim_ssf_data <- tibble(crp_binary = c(rep_len(0, outlen), rep_len(1, outlen)), log_distance_to_street = c(sim.streets, sim.streets), log_sl_ = median(ssf_data$log_sl_), step_id_ = 1, Haukos_SatID = 1) 
# 
# sim_ssf_data %>% 
#   predict(model_figure, newdata = ., type = "link", re.form = NA, se.fit = TRUE) %>%
#   as.data.frame() %>%
#   mutate(predicted.fit = plogis(fit)) ->
#   sim_crp_street
# 
# sim_ssf_data %>%
#   cbind(sim_crp_street) ->
#   sim_ssf_data
# 
# 
# ggplot(sim_ssf_data, mapping = aes(x = log_distance_to_street, y = predicted.fit, color = as.factor(crp_binary))) +
#   geom_line() +
#   geom_ribbon(aes(ymin = predicted.fit - 1.96 * se.fit, ymax = if_else(predicted.fit + 1.96 * se.fit > 1, 1, predicted.fit + 1.96 * se.fit), alpha = 0.3, fill = as.factor(crp_binary))) +
#   ylim(c(0,1)) +
#   facet_grid(cols = vars(crp_binary)) +
#   ylab("Probability of Use") +
#   xlab("Log(Distance to Street)") +
#   theme_bw()

```

The above graph isn't very informative. Instead, let's try and plot the effect sizes for the top 5 models (beta estimates and 95% confidence intervals from the individual models)
```{r}
coef_1 <- broom::tidy(muff.model.set[[23]])
coef_2 <- broom::tidy(muff.model.set[[17]])
coef_3 <- broom::tidy(muff.model.set[[7]])
coef_4 <- broom::tidy(muff.model.set[[6]])
coef_5 <- broom::tidy(muff.model.set[[3]])

coef <- bind_rows(coef_1, coef_2, coef_3, coef_4, coef_5) %>%
  filter(effect == "fixed",
         term != "log_sl_z") %>%
  mutate(name = recode(term,
                       "crp_binary_z" = "CRP at endpoint",
                       "log_distance_to_street_z" = "Distance to street",
                       "prop_path_crp_z" = "CRP grassland\nalong path",
                       "prop_path_grass_z" = "Non-CRP grassland\nalong path",
                       "prop_path_crop_z" = "Cropland\nalong path")) %>%
  mutate(name = factor(name, levels = c("CRP at endpoint", "Distance to street", "CRP grassland\nalong path", "Non-CRP grassland\nalong path", "Cropland\nalong path"), ordered = TRUE))
```

Beta coefficient plot of the 5 most informative models, in order. Don't currently plan to use this in the manuscript 
```{r}
coef %>%
  ggplot(aes(x = name, y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(size = 1.5) +
  theme_bw() +
  geom_errorbar(aes(ymax = estimate + std.error*1.96, ymin = estimate - std.error*1.96), width = 0.3, size = 0.7) +
  labs(x = "Parameters", y = "Beta estimates") +
  scale_y_continuous(breaks = c(-0.3, -0.2, -0.1, 0.0, 0.1, 0.2, 0.3), limits = c(-0.33, 0.33))
```

Doing the above again, but with all models
```{r fig.height=5, fig.width=7.5}
coef_full <- muff.model.set %>%
  map(broom::tidy) %>%
  do.call(bind_rows, .) %>%
  filter(effect == "fixed",
         term != "log_sl_z",
         term != "(Intercept)") %>%
  mutate(category = ifelse(str_detect(term, "prop_path"), "Land cover along step", NA)) %>%
  mutate(category = ifelse(str_detect(term, "cross"), "Obstacle\ncrossing", category)) %>%
  mutate(category = ifelse(str_detect(term, "log_distance"), "Distance from\nendpoint to obstacle", category)) %>%
  mutate(category = ifelse(str_detect(term, "binary"), "Land cover\nat endpoint", category)) %>%
  mutate(category = factor(category, levels = c("Land cover along step", "Land cover\nat endpoint", "Obstacle\ncrossing", "Distance from\nendpoint to obstacle"), ordered = TRUE)) %>%
  mutate(name = recode(term, 
                       "prop_path_crop_z" = "Cropland",
                       "prop_path_shrub_z" = "Shrubland",
                       "prop_path_dev_open_z" = "Developed (open)",
                       "prop_path_grass_z" = "Non-CRP grassland",
                       "prop_path_crp_z" = "CRP grassland",
                       "prop_path_forest_z" = "Forest",
                       "prop_path_barren_z" = "Barren",
                       "prop_path_water_z" = "Water",
                       "prop_path_wetland_z" = "Wetland",
                       "prop_path_dev_lmh_z" = "Developed (low/medium/high)",
                       "cross_secondary_z" = "Secondary road",
                       "cross_street_z" = "Street",
                       "cross_transmission_line_z" = "Transmission line",
                       "log_distance_to_secondary_z" = "Secondary road",
                       "log_distance_to_street_z" = "Street",
                       "log_distance_to_well_z" = "Oil/gas well",
                       "log_distance_to_transmission_line_z" = "Transmission line",
                       "shrub_binary_z" = "Shrubland",
                       "dev_open_binary_z" = "Developed (open)",
                       "grass_binary_z" = "Non-CRP grassland",
                       "crp_binary_z" = "CRP grassland"))
  

# I'm struggling with the interpretation here- log distance to well has the largest effect size by some margin, but is the 8th best model. Meanwhile, the "best" model barely has a significant p-value.

# Land cover at endpoint (CRP) is most informative, presumably because this distinction occurs frequently in the dataset. However, when potential steps fall near obstacles (oil/gas well, roads, secondary power lines)

#
# Note- take out 
#

coef_full %>%
  ggplot(aes(x = name, y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(size = 1.5) +
  theme_bw() +
  geom_errorbar(aes(ymax = estimate + std.error*1.96, ymin = estimate - std.error*1.96), width = 0.3, size = 0.7) +
  labs(x = "Parameters", y = "Beta estimates") + 
  theme(axis.text.x = element_text(angle = 90,hjust=0.95,vjust=0.2)) +
  facet_grid(cols = vars(category), scales = "free_x", space = "free_x")
  #scale_y_continuous(breaks = c(-0.3, -0.2, -0.1, 0.0, 0.1, 0.2, 0.3), limits = c(-0.33, 0.33))
ggsave(filename = "issf_betas_full.tif", device = "tiff", height = 5, width = 7.5)
```


cor(x = ssf_data$crp_binary, y = ssf_data$prop_path_crp)
moderate positive correlation between crp at endpoint and along path (0.6285474)

cor(x = ssf_data$crp_binary, y = ssf_data$log_distance_to_street)
very little correlation between crp at endpoint and distance to street

Calculate how the likelihood of taking a step changes when it does/doesn't end in CRP
Following: https://stackoverflow.com/questions/41384075/r-calculate-and-interpret-odds-ratio-in-logistic-regression
```{r}
crp_nontransformed_model <- glmmTMB(case_ ~ -1 + crp_binary + log_sl_ + (1|step_id_) + (0 + crp_binary|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1)))) %>% 
  fit_glmm_constant_variance() #Build a model

odds_ratio <- broom::tidy(crp_nontransformed_model) %>% 
  filter(term == "crp_binary") %>% 
  pull(estimate) %>% 
  exp()

#This below is probably not correct
#prob_stepping_in_crp <- odds_ratio/(1+odds_ratio)
```

