---
title: "R Notebook"
output: html_notebook
---
Starts after line 38 in issf_glmtmb_analysis, using the results from workspace_12_13_22.RData

Suites:
- Land cover at endpoint (LCAE)
- Land cover along step (LCAS)
- Distance from endpoint to obstacle (DFETO)
- Obstacle crossing (OC)

```{r}
LCAE.candidate.set <- list()

LCAE.candidate.set[[1]] <- glmmTMB(case_ ~ -1 + (1|step_id_), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA)))) #Null

LCAE.candidate.set[[2]] <- glmmTMB(case_ ~ -1 + log_sl_z + (1|step_id_), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA)))) #step length

LCAE.candidate.set[[3]] <- glmmTMB(case_ ~ -1 + shrub_binary_z + log_sl_z + (1|step_id_) + (0 + shrub_binary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

LCAE.candidate.set[[4]] <- glmmTMB(case_ ~ -1 + dev_open_binary_z + log_sl_z + (1|step_id_) + (0 + dev_open_binary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

LCAE.candidate.set[[5]] <- glmmTMB(case_ ~ -1 + grass_binary_z + log_sl_z + (1|step_id_) + (0 + grass_binary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

LCAE.candidate.set[[6]] <- glmmTMB(case_ ~ -1 + crp_binary_z + log_sl_z + (1|step_id_) + (0 + crp_binary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

LCAE.candidate.names <- c("null", "log_sl", "shrub", "dev_open", "grass", "crp")
```

```{r}
LCAS.candidate.set <- list()

LCAS.candidate.set[[1]] <- glmmTMB(case_ ~ -1 + (1|step_id_), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA)))) #Null

LCAS.candidate.set[[2]] <- glmmTMB(case_ ~ -1 + log_sl_z + (1|step_id_), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA)))) #step length

LCAS.candidate.set[[3]] <- glmmTMB(case_ ~ -1 + prop_path_shrub_z + log_sl_z + (1|step_id_) + (0 + prop_path_shrub_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

LCAS.candidate.set[[4]] <- glmmTMB(case_ ~ -1 + prop_path_dev_open_z + log_sl_z + (1|step_id_) + (0 + prop_path_dev_open_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

LCAS.candidate.set[[5]] <- glmmTMB(case_ ~ -1 + prop_path_grass_z + log_sl_z + (1|step_id_) + (0 + prop_path_grass_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

LCAS.candidate.set[[6]] <- glmmTMB(case_ ~ -1 + prop_path_crp_z + log_sl_z + (1|step_id_) + (0 + prop_path_crp_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

LCAS.candidate.names <- c("null", "log_sl", "shrub", "dev_open", "grass", "crp")
```

```{r}
DFETO.candidate.set <- list()

DFETO.candidate.set[[1]] <- glmmTMB(case_ ~ -1 + (1|step_id_), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA)))) #Null

DFETO.candidate.set[[2]] <- glmmTMB(case_ ~ -1 + log_sl_z + (1|step_id_), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA)))) #step length

DFETO.candidate.set[[3]] <- glmmTMB(case_ ~ -1 + log_distance_to_secondary_z + log_sl_z + (1|step_id_) + (0 + log_distance_to_secondary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

DFETO.candidate.set[[4]] <- glmmTMB(case_ ~ -1 + log_distance_to_street_z + log_sl_z + (1|step_id_) + (0 + log_distance_to_street_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

DFETO.candidate.set[[5]] <- glmmTMB(case_ ~ -1 + log_distance_to_well_z + log_sl_z + (1|step_id_) + (0 + log_distance_to_well_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

DFETO.candidate.set[[6]] <- glmmTMB(case_ ~ -1 + log_distance_to_transmission_line_z + log_sl_z + (1|step_id_) + (0 + log_distance_to_transmission_line_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

DFETO.candidate.names <- c("null", "log_sl", "highway", "street", "well", "transmission_line")
```

```{r}
OC.candidate.set <- list()

OC.candidate.set[[1]] <- glmmTMB(case_ ~ -1 + (1|step_id_), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA)))) #Null

OC.candidate.set[[2]] <- glmmTMB(case_ ~ -1 + log_sl_z + (1|step_id_), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA)))) #step length

OC.candidate.set[[3]] <- glmmTMB(case_ ~ -1 + cross_street_z + log_sl_z + (1|step_id_) + (0 + cross_street_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

OC.candidate.set[[4]] <- glmmTMB(case_ ~ -1 + cross_secondary_z + log_sl_z + (1|step_id_) + (0 + cross_secondary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

OC.candidate.set[[5]] <- glmmTMB(case_ ~ -1 + cross_transmission_line_z + log_sl_z + (1|step_id_) + (0 + cross_transmission_line_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

OC.candidate.names <- c("null", "log_sl", "street", "highway", "transmission_line")
```

## Fitting the candidate models

Set the variance as constant at 1000 and fit the model
```{r}
fit_glmm_constant_variance <- function(x){
  #Fix the standard deviation of the first random term, which is the (1|step_id_) component in the above model equation
  #set the variance to 1000 by setting its log (sd) to log(1000)
  x$parameters$theta[1] <- log(1e3)
  
  #We need to tell glmmTMB not to change the variance by setting it to NA:
  #now set when I define each individual function, so this isn't necessary here
  #x$mapArg <- list(theta=factor(c(NA)))
  
  #print so I know which model to attribute error code to
  x$fr %>% names() %>% print()
  
  #Then fit the model and look at the results:
  glmmTMB:::fitTMB(x) %>% 
    return()
}

LCAE.candidate.set %>%
  map(.f = fit_glmm_constant_variance) ->
  LCAE.model.set

LCAS.candidate.set %>%
  map(.f = fit_glmm_constant_variance) ->
  LCAS.model.set

DFETO.candidate.set %>%
  map(.f = fit_glmm_constant_variance) ->
  DFETO.model.set

OC.candidate.set %>%
  map(.f = fit_glmm_constant_variance) ->
  OC.model.set
```

Land cover at endpoint results
```{r}
aictab(cand.set = LCAE.model.set, modnames = LCAE.candidate.names)
```

Land cover along step results
```{r}
aictab(cand.set = LCAS.model.set, modnames = LCAS.candidate.names)
```

Distance from endpoint to obstacle results
```{r}
aictab(cand.set = DFETO.model.set, modnames = DFETO.candidate.names)
```

Obstacle crossing results
```{r}
aictab(cand.set = OC.model.set, modnames = OC.candidate.names)
```

Ensemble suite using the best models: LCAE crp, LCAS crp, DFETO street
```{r}
ensemble.candidate.set <- list()

ensemble.candidate.set[[1]] <- glmmTMB(case_ ~ -1 + (1|step_id_), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA)))) #Null

ensemble.candidate.set[[2]] <- glmmTMB(case_ ~ -1 + log_sl_z + (1|step_id_), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA)))) #step length

ensemble.candidate.set[[3]] <- glmmTMB(case_ ~ -1 + crp_binary_z + log_sl_z + (1|step_id_) + (0 + crp_binary_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

ensemble.candidate.set[[4]] <- glmmTMB(case_ ~ -1 + prop_path_crp_z + log_sl_z + (1|step_id_) + (0 + prop_path_crp_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

ensemble.candidate.set[[5]] <- glmmTMB(case_ ~ -1 + log_distance_to_street_z + log_sl_z + (1|step_id_) + (0 + log_distance_to_street_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1))))

# ensemble.candidate.set[[6]] <- glmmTMB(case_ ~ -1 + crp_binary_z + prop_path_crp_z + log_distance_to_street_z + log_sl_z + (1|step_id_) + (0 + crp_binary_z|Haukos_SatID) + (0 + prop_path_crp_z|Haukos_SatID) + (0 + log_distance_to_street_z|Haukos_SatID), family = poisson, data = ssf_data, doFit = FALSE, map = list(theta=factor(c(NA, 1, 2, 3)))) #full


ensemble.candidate.names <- c("null", "log_sl", "LCAE crp", "LCAS crp", "DFETO street") #, "full"

ensemble.candidate.set %>%
  map(.f = fit_glmm_constant_variance) ->
  ensemble.model.set
```

```{r}
aictab(cand.set = ensemble.model.set, modnames = ensemble.candidate.names)
```

