---
title: "R Notebook"
output: html_notebook
---
Running in R 4.0.5

```{r}
library(raster)
library(tidyverse)
library(googlesheets4)
library(sf)
library(there) #fixed; some code is used to it being broken
library(readr)
library(lubridate)
library(mapview)
library(amt)
library(adehabitatLT)
library(hms)
library(ggplot2)
library(parallel)
library(AICcmodavg)
```

## Importing and cleaning data

Sat locations
Taking only those birds with at least 140 GPS fixes (two weeks on the ground)
```{r}
sat_locations <- read_csv(here_file("used_available_toolshare", "sat_locations_clean", "sat_locations_2019-11-10.csv"), col_types = cols(Haukos_SatID = col_character()))

sat_locations$Haukos_SatID %>%
  str_pad(width = 11, side = "right", pad = "0") ->
  sat_locations$Haukos_SatID

sat_locations %>%
  group_by(Haukos_SatID) %>%
  mutate(count = n()) %>%
  filter(count > 139) %>%
  dplyr::select(-count) %>%
  ungroup() ->
  sat_locations
```

Banding info for other covariates
```{r}
banding_2019 <- read_sheet("1Ys9wnY8GV-kEO9q1hXWbvU3RGAzDe8L2lqWIX3wqQMw", "2019")
banding_2018 <- read_sheet("1Ys9wnY8GV-kEO9q1hXWbvU3RGAzDe8L2lqWIX3wqQMw", "2018")
banding_2017 <- read_sheet("1Ys9wnY8GV-kEO9q1hXWbvU3RGAzDe8L2lqWIX3wqQMw", "2017")
banding_2016 <- read_sheet("1Ys9wnY8GV-kEO9q1hXWbvU3RGAzDe8L2lqWIX3wqQMw", "2016")

banding <- rbind(banding_2016, banding_2017, banding_2018, banding_2019)

banding %>%
  filter(!is.na(BandNumber)) ->
  banding
```

Attaching capture data
```{r}
sat_locations %>%
  mutate(Date = mdy(Date),
         DateTime = mdy_hm(DateTime),
         Transmitter = "SAT-PTT") %>%
  left_join(dplyr::select(banding, Haukos_SatID, BandNumber, Sex, ReleaseSite, ReleaseLocation, ReleaseYear, HandlingDate)) %>%
  dplyr::select(BandNumber, Haukos_SatID, Date, DateTime, Latitude, Longitude, Sex, ReleaseSite, ReleaseLocation, ReleaseYear, HandlingDate) ->
  sat_locations
```

Importing dispersal settlement dates so I can narrow down to dispersal locations
```{r}
st_read(dsn = here_file("Projects", "Dispersal_paper", "settlement_designations"), layer = "settlement_locations_2018") %>%
  st_drop_geometry() %>%
  transmute(Haukos_SatID = Hks_SID,
         settlement_date = DateTim) ->
  settlement_2018

st_read(dsn = here_file("Projects", "Dispersal_paper", "settlement_designations"), layer = "settlement_locations_2019") %>%
  st_drop_geometry() %>%
transmute(Haukos_SatID = Hks_SID,
         settlement_date = DateTim) ->
  settlement_2019

settlement_dates <- rbind(settlement_2018, settlement_2019)
```

A large proportion of those missing dispersal end dates have trajectories ending before 7/31, and were discarded as early deaths. I manually added those in as a settlement date of Jan 1 2021. Removed 7 with an NA settlement date, where the end of the dispersal movement could not be determined

EXCEPT some large portion of these birds settle before 7/31 and then die. I need to figure out how to discern those locations. Probably means revisiting the original algorithm?

^Which I did

Trajectories where all points are dispersing: set end date to 1/1/2021
Trajectories which we can't use, either because it doesn't fit the pattern or the algorithm couldn't determine a breaking point: set end date to 1/1/1970

2019.172607 died too early (5 days on ground) and so is removed here as well
As did:
2018.172610
2018.172626
2018.172639
2018.172641

One excluded (2019.172591) because they didn't reach an end point within the timeframe of the summer

```{r}
textConnection(
"Haukos_SatID, settlement_date
2018.172603, 1/1/1970
2018.172630, 1/1/2021
2018.172637, 1/1/1970
2018.172640, 1/1/2021
2018.172642, 1/1/1970
2018.172593, 1/1/2021
2018.172596, 1/1/2021
2018.172608, 1/1/2021
2018.172609, 1/1/2021
2018.172611, 1/1/2021
2018.172621, 1/1/2021
2019.172587, 1/1/2021
2019.172598, 1/1/2021
2019.172600, 1/1/2021
2019.172615, 1/1/2021
2019.174277, 1/1/2021
2019.172597, 1/1/2021
2019.172602, 1/1/1970
2019.172611, 1/1/1970
2019.172616, 1/1/1970
2019.172636a, 1/1/1970
2019.174255, 1/1/2021
2019.174256, 1/1/1970
2019.174268, 1/1/2021
2019.172591, 1/1/1970
2019.172604, 1/1/1970
2019.172610, 1/1/1970
2019.174272, 1/1/1970
2019.174280, 1/1/1970
2019.172607, 1/1/1970
2018.172610, 1/1/1970
2018.172626, 1/1/1970
2018.172639, 1/1/1970
2018.172641, 1/1/1970
") %>% 
  read.csv(strip.white = TRUE, colClasses = c("character", "character")) %>%
  mutate(settlement_date = mdy(settlement_date)) ->
  settlement_dates_appended
  
```

Using settlement dates to constrain locations to those before settlement occurs
```{r}
settlement_dates <- rbind(settlement_dates, settlement_dates_appended)

sat_locations %>%
  left_join(settlement_dates) %>%
  filter(Date < settlement_date) ->
  sat_locations
```

Cleaning the remaining locations

Mortality
```{r}
mort_ks <- read_sheet("1WnEf_bP3W_cDHcIVL6bBAnnBMkLRDlpEi0mamg_2TU8", "KS Mortality")
mort_co <- read_sheet("1WnEf_bP3W_cDHcIVL6bBAnnBMkLRDlpEi0mamg_2TU8", "CO Mortality")

mort <- rbind(mort_ks, mort_co)
```

Get rid of all locations on and prior to the release date,
Get rid of all locations on and following the mort date
```{r}
sat_locations %>%
  filter(Date > HandlingDate) %>%
  left_join(dplyr::select(mort, BandNumber, Mortality_Detection_Date)) %>%
  mutate(Mortality_Detection_Date = as_date(Mortality_Detection_Date)) %>%
  mutate(Mortality_Detection_Date = replace_na(Mortality_Detection_Date, mdy("1/1/2021"))) %>%
  filter(Date < Mortality_Detection_Date) -> 
  sat_locations
```

Checked here: the dates of all steps are reasonable and fall within summers of 2018 and 2019

## Resampling to 1 location per day

New code: resampling these to a single location per bird per day, to remove variation in step length between hours

This is due to a) this not working otherwise b) Erik making the valid point that habitat selection changes between behavioral modes, and there are multiple behavioral modes per day

Keeping with all locations at 23:00 GMT, as 22:00 to 23:00 has the lowest movement speed of any time period. I'm basically measuring steps from night roost to night roost
```{r}
sat_locations %>%
  filter(hour(DateTime) == 23) -> sat_locations
```

## Preparing a step selection function

Start by reading into the amt track format
```{r}
sat_locations %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) %>%
  st_transform(5070) %>%
  st_coordinates() ->
  coords_5070

sat_locations <- cbind(sat_locations, coords_5070)

sat_locations %>%
  dplyr::select(BandNumber, DateTime) %>%
  duplicated() -> 
  duplicated_indicies

sat_locations[!duplicated_indicies,] -> sat_locations
  
sat_locations %>%
  make_track(.x = X, .y = Y, .t = DateTime, bird = BandNumber, crs = sp::CRS("+init=epsg:5070"), all_cols = TRUE)  -> 
  lpc_track

```

Nesting the dataframe into a list within another dataframe
```{r}
lpc_movements <- lpc_track %>% nest(tracks = -"BandNumber")
```

Converting to steps
```{r}
lpc_movements %>%
  mutate(steps = map(tracks, function(x)
    x %>%
      steps(keep_cols = 'start'))) ->
  lpc_movements
```

Calculating velocity for the steps
```{r}
lpc_movements %>%
  mutate(steps = map(steps, function(x)
    x %>%
      mutate(velocity = sl_/as.numeric(dt_)))) -> 
  lpc_movements
```

Now for the tricky part: I'm going to create a separate steps_xyt object for each time, so that I can take advantage of time-specific step length distributions
Each column is named after the UTC starting time for the step

Update: This did NOT WORK. Instead using an integrated step selection function to account for the interaction between time and step length during the conditional logistic regression phase, as specified in the vignette at https://arxiv.org/pdf/1805.03227.pdf

Update: This also didn't work, so I just did a daily resampling to remove the hourly variation

```{r}
#Removing individuals with 1 step, as this causes errors
lpc_movements %>%
  mutate(num_steps = map(steps, nrow)) %>%
  filter(num_steps > 1) ->
  lpc_movements

lpc_movements %>%
  mutate(steps_full = map(steps, function(x)
    x %>%
      random_steps(n = 9))) -> 
  lpc_movements
```

## Extract covariates to each step
Reading those covariates in

Key for landcov analyses:
0 No data
1 Crop
2 Forest/Woody wetland
3 Shrubland
4 Barren
6 Water
7 Wetland
8 Developed (open space, roads)
9 Developed (low, mid, high)
10 Grassland/Pasture
11 CRP

```{r}
dist_secondary <- raster(here_file("Projects", "Dispersal_paper", "distance_to_secondary.tif"))

dist_street <- raster(here_file("Projects", "Dispersal_paper", "distance_to_street.tif"))

dist_well <- raster(here_file("Projects", "Dispersal_paper", "distance_to_well.tif"))

dist_transmission_line <- raster(here_file("Projects", "Dispersal_paper", "distance_to_transmission_line.tif"))

cross_secondary <- raster(here_file("Projects", "Dispersal_paper", "crossing_secondary.tif"))

cross_street <- raster(here_file("Projects", "Dispersal_paper", "crossing_street.tif"))

cross_transmission_line <- raster(here_file("Projects", "Dispersal_paper", "crossing_transmission_line.tif"))

landcov_2018 <- raster(here_file("Projects", "Dispersal_paper", "landcover_classification", "dispersing_landcov_2018_5070.tif")) #"Projects", "Dispersal_corridors",
names(landcov_2018) <- "landcov_18"

landcov_2019 <- raster(here_file("Projects", "Dispersal_paper", "landcover_classification", "dispersing_landcov_2019_5070.tif"))
names(landcov_2019) <- "landcov_19" #"Projects", "Dispersal_corridors",

patchsize_2018 <- raster(here_file("Projects", "Dispersal_paper", "landcover_classification", "dispersing_landcov_2018_patch_size.tif")) #"Projects", "Dispersal_corridors",
names(patchsize_2018) <- "patch_size_18"

patchsize_2019 <- raster(here_file("Projects", "Dispersal_paper", "landcover_classification", "dispersing_landcov_2019_patch_size.tif"))#"Projects", "Dispersal_corridors",
names(patchsize_2019) <- "patch_size_19"
```

Join the covariates to the endpoint of each step
```{r}
lpc_movements %>%
  mutate(steps_covariates = map(steps_full, function(x)
    x %>%
      extract_covariates(dist_secondary))) -> 
  lpc_movements

lpc_movements %>%
  mutate(steps_covariates = map(steps_covariates, function(x)
    x %>%
      extract_covariates(dist_street))) -> 
  lpc_movements

lpc_movements %>%
  mutate(steps_covariates = map(steps_covariates, function(x)
    x %>%
      extract_covariates(dist_well))) -> 
  lpc_movements

lpc_movements %>%
  mutate(steps_covariates = map(steps_covariates, function(x)
    x %>%
      extract_covariates(dist_transmission_line))) -> 
  lpc_movements

save(lpc_movements, file = "lpc_movements.RData")
```

Using parallel processing to determine when LPC cross obstacles
First: crossing secondary roads
```{r}
numCores <- detectCores() - 2
numCores

cl <- makeCluster(numCores)

lpc_movements$steps_covariates -> steps_covariates

clusterExport(cl, "steps_covariates")
clusterExport(cl, "cross_secondary")

clusterEvalQ(cl, {
  library(tidyverse)
  library(amt)
})

clusterApplyLB(cl, steps_covariates, function(x)
    x %>%
      extract_covariates_along(cross_secondary)) -> along_secondary

save(along_secondary, file = "along_secondary.RData")
stopCluster(cl)
```

Second: crossing streets
```{r}
numCores <- detectCores() - 1
numCores

cl <- makeCluster(numCores)

clusterExport(cl, "steps_covariates")
clusterExport(cl, "cross_street")

clusterEvalQ(cl, {
  library(tidyverse)
  library(amt)
})

clusterApplyLB(cl, steps_covariates, function(x)
    x %>%
      extract_covariates_along(cross_street)) -> along_street

save(along_street, file = "along_street.RData")
stopCluster(cl)
```

Third: crossing transmission lines
```{r}
numCores <- detectCores() - 1
numCores

cl <- makeCluster(numCores)

clusterExport(cl, "steps_covariates")
clusterExport(cl, "cross_transmission_line")

clusterEvalQ(cl, {
  library(tidyverse)
  library(amt)
})

clusterApplyLB(cl, steps_covariates, function(x)
    x %>%
      extract_covariates_along(cross_transmission_line)) -> along_transmission_line

save(along_transmission_line, file = "along_transmission_line.RData")
stopCluster(cl)
```

Fourth: identity of 2018 landcover pixels crossed
```{r}
numCores <- detectCores() - 1
numCores

cl <- makeCluster(numCores)

clusterExport(cl, "steps_covariates")
clusterExport(cl, "landcov_2018")

clusterEvalQ(cl, {
  library(tidyverse)
  library(amt)
})

clusterApplyLB(cl, steps_covariates, function(x)
    x %>%
      extract_covariates_along(landcov_2018)) -> along_landcov_18

save(along_landcov_18, file = "along_landcov_18.RData")
stopCluster(cl)
```

Fifth: identity of 2019 landcover pixels crossed
```{r}
numCores <- detectCores() - 1
numCores

cl <- makeCluster(numCores)

clusterExport(cl, "steps_covariates")
clusterExport(cl, "landcov_2019")

clusterEvalQ(cl, {
  library(tidyverse)
  library(amt)
})

clusterApplyLB(cl, steps_covariates, function(x)
    x %>%
      extract_covariates_along(landcov_2019)) -> along_landcov_19

save(along_landcov_19, file = "along_landcov_19.RData")
stopCluster(cl)
```

Now summarizing the barrier/landcover crossing data and attaching it as a covariate
```{r}
along_secondary %>% map(function(x){
  map(x, function(i)
    any(i == 1) %>%
    if_else(1,0))
}) -> along_secondary_binary

map2(steps_covariates, along_secondary_binary, function(x, y)
  x %>%
    mutate(cross_secondary = unlist(y))) ->
  steps_covariates

along_street %>% map(function(x){
  map(x, function(i)
    any(i == 1) %>%
    if_else(1,0))
}) -> along_street_binary

map2(steps_covariates, along_street_binary, function(x, y)
  x %>%
    mutate(cross_street = unlist(y))) ->
  steps_covariates

along_transmission_line %>% map(function(x){
  map(x, function(i)
    any(i == 1) %>%
    if_else(1,0))
}) -> along_transmission_line_binary

map2(steps_covariates, along_transmission_line_binary, function(x, y)
  x %>%
    mutate(cross_transmission_line = unlist(y))) ->
  steps_covariates
```

Attaching proportions of each landcover type from 2018
```{r}
along_landcov_18 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 1, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_crop_18

map2(steps_covariates, prop_path_crop_18, function(x, y)
  x %>%
    mutate(prop_path_crop_18 = unlist(y))) ->
  steps_covariates

along_landcov_18 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 2, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_forest_18

map2(steps_covariates, prop_path_forest_18, function(x, y)
  x %>%
    mutate(prop_path_forest_18 = unlist(y))) ->
  steps_covariates

along_landcov_18 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 3, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_shrub_18

map2(steps_covariates, prop_path_shrub_18, function(x, y)
  x %>%
    mutate(prop_path_shrub_18 = unlist(y))) ->
  steps_covariates

along_landcov_18 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 4, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_barren_18

map2(steps_covariates, prop_path_barren_18, function(x, y)
  x %>%
    mutate(prop_path_barren_18 = unlist(y))) ->
  steps_covariates

along_landcov_18 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 6, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_water_18

map2(steps_covariates, prop_path_water_18, function(x, y)
  x %>%
    mutate(prop_path_water_18 = unlist(y))) ->
  steps_covariates

along_landcov_18 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 7, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_wetland_18

map2(steps_covariates, prop_path_wetland_18, function(x, y)
  x %>%
    mutate(prop_path_wetland_18 = unlist(y))) ->
  steps_covariates

along_landcov_18 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 8, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_dev_open_18

map2(steps_covariates, prop_path_dev_open_18, function(x, y)
  x %>%
    mutate(prop_path_dev_open_18 = unlist(y))) ->
  steps_covariates

along_landcov_18 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 9, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_dev_lmh_18

map2(steps_covariates, prop_path_dev_lmh_18, function(x, y)
  x %>%
    mutate(prop_path_dev_lmh_18 = unlist(y))) ->
  steps_covariates

along_landcov_18 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 10, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_grass_18

map2(steps_covariates, prop_path_grass_18, function(x, y)
  x %>%
    mutate(prop_path_grass_18 = unlist(y))) ->
  steps_covariates

along_landcov_18 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 11, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_crp_18

map2(steps_covariates, prop_path_crp_18, function(x, y)
  x %>%
    mutate(prop_path_crp_18 = unlist(y))) ->
  steps_covariates

```

And now 2019
```{r}
along_landcov_19 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 1, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_crop_19

map2(steps_covariates, prop_path_crop_19, function(x, y)
  x %>%
    mutate(prop_path_crop_19 = unlist(y))) ->
  steps_covariates

along_landcov_19 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 2, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_forest_19

map2(steps_covariates, prop_path_forest_19, function(x, y)
  x %>%
    mutate(prop_path_forest_19 = unlist(y))) ->
  steps_covariates

along_landcov_19 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 3, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_shrub_19

map2(steps_covariates, prop_path_shrub_19, function(x, y)
  x %>%
    mutate(prop_path_shrub_19 = unlist(y))) ->
  steps_covariates

along_landcov_19 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 4, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_barren_19

map2(steps_covariates, prop_path_barren_19, function(x, y)
  x %>%
    mutate(prop_path_barren_19 = unlist(y))) ->
  steps_covariates

along_landcov_19 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 6, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_water_19

map2(steps_covariates, prop_path_water_19, function(x, y)
  x %>%
    mutate(prop_path_water_19 = unlist(y))) ->
  steps_covariates

along_landcov_19 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 7, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_wetland_19

map2(steps_covariates, prop_path_wetland_19, function(x, y)
  x %>%
    mutate(prop_path_wetland_19 = unlist(y))) ->
  steps_covariates

along_landcov_19 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 8, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_dev_open_19

map2(steps_covariates, prop_path_dev_open_19, function(x, y)
  x %>%
    mutate(prop_path_dev_open_19 = unlist(y))) ->
  steps_covariates

along_landcov_19 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 9, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_dev_lmh_19

map2(steps_covariates, prop_path_dev_lmh_19, function(x, y)
  x %>%
    mutate(prop_path_dev_lmh_19 = unlist(y))) ->
  steps_covariates

along_landcov_19 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 10, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_grass_19

map2(steps_covariates, prop_path_grass_19, function(x, y)
  x %>%
    mutate(prop_path_grass_19 = unlist(y))) ->
  steps_covariates

along_landcov_19 %>% map(function(x){
  map(x, function(i)
    i %>%
      map(.f = function(x){
        if_else(x == 11, 1, 0)
        }) %>%
      unlist() %>%
      mean()
    )
}) -> prop_path_crp_19

map2(steps_covariates, prop_path_crp_19, function(x, y)
  x %>%
    mutate(prop_path_crp_19 = unlist(y))) ->
  steps_covariates

```

Extracting a few last sets of land cover covariates to endpoints
```{r}
map(steps_covariates, function(x)
    x %>%
      extract_covariates(landcov_2018)) -> 
  steps_covariates

map(steps_covariates, function(x)
    x %>%
      extract_covariates(landcov_2019)) -> 
  steps_covariates
```

and patch size covariates to endpoints
```{r}
map(steps_covariates, function(x)
    x %>%
      extract_covariates(patchsize_2018)) -> 
  steps_covariates

map(steps_covariates, function(x)
    x %>%
      extract_covariates(patchsize_2019)) -> 
  steps_covariates
```

And finally bringing steps_covariates back into the lpc_movements df
```{r}
lpc_movements$steps_covariates_full <- steps_covariates
```

## Cleaning steps & covariates prior to the issf

Kicking out any steps which end outside the study area
```{r}
counties_interest <- raster(here_file("Projects", "Dispersal_paper", "counties_of_interest.tif"))#
names(counties_interest) <- "county"

lpc_movements %>%
  mutate(steps_covariates_full = map(steps_covariates_full, function(x)
    x %>%
      extract_covariates(counties_interest))) -> 
  lpc_movements

lpc_movements %>%
  mutate(steps_covariates_full = map(steps_covariates_full, function(x)
    x %>%
      filter(county == 1))) -> 
  lpc_movements
```

Any step lengths that are zero are changed to the minimum step length for that individual. Allows sl_ to be log transformed
```{r}
lpc_movements %>%
  mutate(steps_covariates_full = map(steps_covariates_full, function(x){
    x %>%
      filter(sl_ != 0) %>%
      pull(sl_) %>%
      min(na.rm = TRUE) -> y
    x %>%
      mutate(sl_ = recode(sl_, "0" = y))
  })) -> 
  lpc_movements
```

Any distance covariates that are zero are changed to 1. Allows these covariates to be log transformed
```{r}
lpc_movements %>%
  mutate(steps_covariates_full = map(steps_covariates_full, function(x){
    x %>%
      mutate(distance_to_secondary = recode(distance_to_secondary, "0" = 1)) %>%
      mutate(distance_to_street = recode(distance_to_street, "0" = 1)) %>%
      mutate(distance_to_well = recode(distance_to_well, "0" = 1)) %>%
      mutate(distance_to_transmission_line = recode(distance_to_transmission_line, "0" = 1))
  })) -> 
  lpc_movements
```

transforming step lengths and distance covariates to be normally distributed
```{r}
lpc_movements %>%
  mutate(steps_covariates_full = map(steps_covariates_full, function(x) {
    x %>%
      mutate(log_sl_ = log(sl_)) %>%
      mutate(log_distance_to_secondary = log(distance_to_secondary)) %>%
      mutate(log_distance_to_street = log(distance_to_street)) %>%
      mutate(log_distance_to_well = log(distance_to_well)) %>%
      mutate(log_distance_to_transmission_line = log(distance_to_transmission_line))
})) ->
  lpc_movements
```

Remove steps where dt_ != 1 (multi-day steps due to missed GPS locations) 
```{r}
lpc_movements %>%
  mutate(steps_covariates_full = map(steps_covariates_full, function(x)
    x %>%
      filter(dt_ == 1))) -> 
  lpc_movements
```

Determining which landcov covariates should be used based on the year
```{r}
lpc_movements %>%
  mutate(steps_covariates_full = map(steps_covariates_full, function(x)
    x %>%
      mutate(landcov = as.factor(if_else(ReleaseYear == 2018, landcov_18, landcov_19))) %>%
      mutate(prop_path_crop = as.numeric(if_else(ReleaseYear == 2018, prop_path_crop_18, prop_path_crop_19))) %>%
      mutate(prop_path_forest = as.numeric(if_else(ReleaseYear == 2018, prop_path_forest_18, prop_path_forest_19))) %>%
      mutate(prop_path_shrub = as.numeric(if_else(ReleaseYear == 2018, prop_path_shrub_18, prop_path_shrub_19))) %>%
      mutate(prop_path_barren = as.numeric(if_else(ReleaseYear == 2018, prop_path_barren_18, prop_path_barren_19))) %>%
      mutate(prop_path_water = as.numeric(if_else(ReleaseYear == 2018, prop_path_water_18, prop_path_water_19))) %>%
      mutate(prop_path_wetland = as.numeric(if_else(ReleaseYear == 2018, prop_path_wetland_18, prop_path_wetland_19))) %>%
      mutate(prop_path_dev_open = as.numeric(if_else(ReleaseYear == 2018, prop_path_dev_open_18, prop_path_dev_open_19))) %>%
      mutate(prop_path_dev_lmh = as.numeric(if_else(ReleaseYear == 2018, prop_path_dev_lmh_18, prop_path_dev_lmh_19))) %>%
      mutate(prop_path_grass = as.numeric(if_else(ReleaseYear == 2018, prop_path_grass_18, prop_path_grass_19))) %>%
      mutate(prop_path_crp = as.numeric(if_else(ReleaseYear == 2018, prop_path_crp_18, prop_path_crp_19))) %>%
      mutate(patch_size = as.numeric(if_else(ReleaseYear == 2018, patch_size_18, patch_size_19)))
      )) -> 
  lpc_movements
```

Key for landcov analyses:
0 No data
1 Crop
2 Forest/Woody wetland
3 Shrubland
4 Barren
6 Water
7 Wetland
8 Developed (open space, roads)
9 Developed (low, mid, high)
10 Grassland/Pasture
11 CRP

Need to plot all of these out before I use a final run! Check for a normal distribution

## Folding into a single tibble for analysis

```{r}
lpc_movements %>% #note: the fewer of these I keep in the better the p values get
  pull(steps_covariates_full) %>%
  bind_rows() %>%
  mutate(landcov = as.factor(recode(as.character(landcov), 
                                     '128' = '1', #error causes this landcov
                                     '1' = '1',
                                     '2' = '2',
                                     '3' ='3',
                                     '4' = '4',
                                     '5' = '5',
                                     '6' = '6',
                                     '7' = '7',
                                     '8' = '8',
                                     '9' = '9',
                                     '10' = '10',
                                     '11' = '11'))) %>% 
  mutate(crop_binary = if_else(landcov == 1, 1, 0),
         forest_binary = if_else(landcov == 2, 1, 0),
         shrub_binary = if_else(landcov == 3, 1, 0),
         barren_binary = if_else(landcov == 4, 1, 0),
         water_binary = if_else(landcov == 6, 1, 0),
         wetland_binary = if_else(landcov == 7, 1, 0),
         dev_open_binary = if_else(landcov == 8, 1, 0),
         dev_lmh_binary = if_else(landcov == 9, 1, 0),
         grass_binary = if_else(landcov == 10, 1, 0),
         crp_binary = if_else(landcov == 11, 1, 0),
         ) %>%
  mutate(step_id_ = paste0(step_id_, "_", Haukos_SatID)) -> 
  ssf_data
```

Save image for future analysis
```{r}
save.image(file = "workspace_XX_XX_XX.RData")
```

